\plaintitle

### 2016-09-20 Paper/figures/figures.tex
Comments for remainder of draft.

\newcommand{\outlierFig}{
  \begin{figure}
  \centering
  \includegraphics[width=.55\columnwidth]{figures/exp4}
  \caption{A cropped example stimulus from experiment 3. We replaced the final 15 points of this dataset with extreme values. The overlaid purple trend line represents a robust OLS fit (ignoring the outlier values), while the overlaid green line represents the fit with all points included. Participants' estimates of trend lines were universally closer to the robust than the non-robust trend; regression by eye tends to underweight outliers compared to OLS.}
  \label{fig:outlier}
  \end{figure}
}

>>>>

\newcommand{\outlierFig}{
  \begin{figure}
  \centering
  \includegraphics[width=.55\columnwidth]{figures/exp4}
  \caption{A cropped example stimulus from experiment 3. We replaced the final 15 points of this dataset with extreme values. The overlaid purple trend line represents a robust OLS fit (ignoring the outlier values), while the overlaid green line represents the fit with all points included. Participants' estimates of trend lines were universally closer to the robust than the non-robust trend; regression by eye tends to underweight outliers compared to OLS. \jeff{Could we include the average trend line produced by participants?}}
  \label{fig:outlier}
  \end{figure}
}

### 2016-09-20 Paper/proceedings.tex
Comments for remainder of draft.

Even if designers include all relevant information, the audience may lack the statistical expertise to interpret these values, or may be misled if the data violate modeling assumptions. As a further complication, the form of visual encoding may influence viewers' inferences. For instance, viewers may be more likely to consider trends with line charts, and to compare individual values with bar charts \cite{zacks1999bars}. Visual design choices can also introduce bias, such as the visual asymmetry of bars causing ``within-the-bar'' bias \cite{newman2012bar}. Designers would benefit from guidance regarding how accurately viewers make trend estimations by eye, and to what degree different visualization types and annotations might bias these estimations.

@@

\section{1. Estimation of Slope is Reliable \\ Across Visualizations}

We designed our first experiment to examine how accurate participants were are estimating the direction (slope, amplitude, or curvature) of trends in bivariate visualizations. We examined a variety of common bivariate visualizations. In addition to linear trends, we examined more complex relationships such as quadratic and trigonometric trends.

We presented participants with a series of bivariate visualizations, and adjusted a slider to fit the perceived trend in the points. Visualizations were one of a scatter plot, linegraph, or area chart (with the filled area below the line). Visualizations consisted of one of three types of trend (linear, quadratic, or trigonometric). For each stimulus, participants adjusted a slider that controlled the slope of a rendered trend line (or, in the case of the quadratic trends, the curvature, or the trigonometric, the positive/negative amplitude).

Participants saw one of each combination of 3 chart types (scatter plot, line chart, or area chart), 8 possible slopes $\beta = \pm \{0.1,0.2,0.4,0.8\}$, 4 bandwidths of Gaussian residuals $\sigma = \{0.05,0.1,0.15,0.2\}$, for a total of 96 stimuli, with an additional 4 validation stimuli otherwise excluded from analysis. Type of trend (linear, quadratic, or trigonometric) was a random factor, with 32 stimuli of each factor level.

@@

We had three hypotheses for the first experiment:
\begin{enumerate}
	\item \textbf{As the bandwidth of the residuals increased, accuracy would decrease}. Increasing the bandwidth of the residuals results in a lower correlation coefficient and higher perceived noise in the bivariate data. Prior work indicates that these related measures correspond to decreased accuracy for aggregate tasks in bivariate visualizations \cite{albers2014task,harrison2014ranking}.
	\item \textbf{More complex relationships would result in lower accuracy}. Quadratic and trigonometric relationships are visually more complex than linear relationships, and often require more complex statistical methods to analyze. We anticipated that estimation of these less familiar relationships would therefore be more difficult than in the linear case.
	\item \textbf{Estimations would be unbiased}. That is, there would be no systematic over- or underestimation of trends.
\end{enumerate}
\subsection{Results}
We performed a three-way analysis of covariance (ANCOVA) of the effect of residual bandwidth, graph type and trend type on unsigned error in estimation of trend lines. We included participant ID and the actual slope of the trend line as covariates.

Our results support our first hypothesis: \textbf{larger residuals reduce accuracy at regression by eye}. Bandwidth of the Gaussian used to generate residuals was a significant main effect ($F(1,4554)=950$), and the trimmed mean of absolute error increased monotonically with this bandwidth, from $7\%$ of the actual slope of the trend line when the bandwidth was $0.05$, to $32\%$ when the bandwidth was $0.20$. Figure \ref{fig:sigma} illustrates these results.

Our results fail to support our second hypothesis: \textbf{there was no statistically significant difference in estimation accuracy among linear, quadratic, or trigonometric trends}. Fit type was only a marginal main effect ($F(2,4554)=2.60$, $p=0.074$), and a post-hoc Tukey's test of Honest Significant Difference (HSD) did not identify any significant pairwise-interactions. The trimmed mean of absolute estimation error of linear fits was $17\%$ of the actual slope, in the middle of the $16\%$ error for quadratic fits, and $18\%$ for trignometric fits, providing evidence that the relative unfamiliarity of non-linear trends does not have a significant (in the ordinary sense) impact on performance.

Our results support our third hypothesis: \textbf{there was no statistically significant bias in estimations}. Participants saw an equal number of stimuli with positive and negative trend lines. If estimates of these trends were unbiased, we would expect the average signed error to be close to zero. The average signed error was $0.0008$, far less than the fidelity of the slider used to input guesses (which had $\Delta=0.01$). A Student's T test failed to support ($p=0.70$) the hypothesis that $mu_{error}\ne0$.

\section{2. Estimation of Intercept is Subject to \\ ``Within The Bar'' Bias}

``Within the bar'' bias is a perceptual bias in bar charts where points contained in the visual area of the glyph of the bar are seen as likelier than points outside of the glyph. Newman \& Scholl \cite{newman2012bar} encountered this bias for a sampling task: ``how likely is this point to have been drawn from the distribution represented by this bar?'' Correll \& Gleicher \cite{correll2014error} likewise encountered this bias for inferential tasks: ``how likely is the population mean to take a particular value, given the sample represented by this bar?'' We believed that this bias would likewise occur in regression tasks in visually asymmetric visualizations such as area charts. The slope estimation task from the previous experiment would not capture this bias, as there is no method for participants to indicate a \emph{uniform} underestimation in trends; decreasing the slope would cause underestimation at the beginning of the plot but not the end, and vice versa. We therefore designed this experiment to measure estimates of the y-intercepts of trends. The within-the-bar bias here would then appear as systematic underestimation of intercept in area charts.

@@

We had one hypothesis for the second experiment:
\begin{enumerate}
	\item \textbf{Area charts would be subject to within the bar bias.} That is, participants would estimate lower values of y-intercepts of trends than with line charts and scatter plots.
\end{enumerate}
\subsection{Results}
We performed a one-way ANCOVA of the effect of graph type on \emph{signed} error. We included residual bandwidth as a covariate, and participant ID as a random factor.

Our results support our first hypothesis: \textbf{participants systematically overestimated the intercept of trends in area charts, but not in scatter plots or linegraphs}. Graph type was a significant main effect on signed error ($F(2,2431)=27$, $p<0.001$). A post-hoc Tukey's HSD confirmed significant differences in error between the area chart and the other two chart types, but not between scatter plots and linegraphs. The trimmed mean of the signed error of estimations made with area charts was an underestimation of $-0.02$, compared to a trimmed mean of $0.002$ for the other two conditions. This underestimation corresponded to unsigned errors more than twice as large in area charts (trimmed mean of $0.04$ for intercepts that ranged from $\{-0.25,0.25\}$, compared to $0.02$ for linegraphs and scatter plots). Figure \ref{fig:outlier} illustrates this result.

\section{3. Estimation of Slope is Robust to Outliers}

\outlierFig
OLS regression operates under the assumption that there is a unimodal, symmetric distribution of residuals surrounding the line of best fit. Extreme outliers violate this assumption, and can result in trend lines that are substantially different from those produced by more robust methods. Visual inspection can accurately identify certain classes of outliers \cite{albers2014task}. Therefore, regression by eye may afford the estimation of both outlier-robust and outlier-sensitive trends. However, cognitive biases such as anchoring (the tendency to overweight the first set of information), availability (the tendency to overweight more recent or extreme information), and the hot-hand fallacy (the tendency to assume that trends of high or low values with continue) can impact how viewers categorize and utilize outliers in their estimations. These biases can impact forecasting tasks~\cite{campbell2009anchoring,ji2001culture}. That is, regression by eye may not uniformly weight outliers, depending on their position in the plot.

@@

We had three hypotheses for the third experiment:
\begin{enumerate}
	\item \textbf{Participant estimation would be closer to an outlier-less robust trend line, than the non-robust OLS trend line.} That is, we assumed that in general, participants would ignore or downweight outliers when performing regression by eye.
	\item \textbf{As the number of outliers increased, estimations would be closer to the non-robust OLS fit.} We speculated on the existence of a ``tipping point'' of outlier density, beyond which participants would avoid robust fits, and interpolate between the robust and OLS fits.
	\item \textbf{Outliers at the end of the chart would result in estimations closer to the OLS fit than outliers in other locations on the plot.} Prior work on cognitive biases in forecasting tasks suggests that viewers may give more credence to more recent outliers (as indicative as new anchors points, or of an emerging ``streak''). We therefore believed that these points would be more heavily weighted, and the resulting estimation closer to the outlier-sensitive OLS line, than outliers ``earlier'' in the plot.
\end{enumerate}
\subsection{Results}

Our results support our first hypothesis. \textbf{Participant estimates were closer to the robust trend line than the trend line that included outliers}. Excluding conditions with zero outliers (and so the two trend lines would be identical), the trimmed mean of unsigned error was over $4$ times higher when calculated as a comparison to the OLS trend line ($0.36$) as opposed to the robust trend line ($0.08$). A Student's T-test confirmed that the error as defined by the OLS line was significantly higher than the robust line ($p<0.001$). However, a Student's T test found a significant difference in unsigned error when compared to the robust trend line between estimates where no outliers were present, and those with any amount of outliers ($M=0.07$ vs. $M=0.08$, $p<0.001$). This indicates that participants are not entirely ignoring outliers, although the small effect size suggestes that participants are giving outliers significantly less weight than OLS regression.

@@

In many cases, designers do not explicitly encoded regression information. In other cases, viewers may not have the statistical or graphical expertise to interpret such information, even when it is supplied. Yet, our results point to regression by eye as a robust and accurate method for estimating trends in bivariate data. Participants from a variety of backgrounds and levels of self-reported graphical and statistical expertise were capable of estimating both the slope and intercept of both linear and non-linear trends. That is, viewers of visualization are largely trustworthy when estimating the relationship between two variables in a plot.

However, this general accuracy at estimation trends is not universal. Area charts are visually asymmetrical: area below the line is filled in, and the area above it as not. This asymmetry creates within the bar bias: a systematic underestimation of the intercept of trends, due to the perceived higher likelihood of points in the filled-in area. Designers hoping to rely on regression by eye should avoid such asymmetries in their bivariate visualizations.

Likewise, viewers do not give the same weight to outliers as certain standard forms of statistical regression such as OLS. For noisy data, this robustness may be beneficial: humans themselves can be relied on to perform filtering operations without explicit guidance. However, in other cases, this insensitivity may result in visual estimations at odds with the statistical modeling underpinning other areas of analysis of a particular dataset. The emergence of new modes and the shifting of means may not result in a subsequent shift in beliefs about trends in data.

@@

Our experimental setup was intentionally simplistic, affording only a single free parameter for each experiment. In actual regression by eye, the viewer must simultaneously engage in multiple types of estimation: choosing a particular type of fit, filtering data, and then estimating the parameters of the chosen model. Errors in any one of these steps could compound, resulting in performance worse than our measures, where many of these decisions are presented as a given.

@@

Finally, we focused on a simple form of regression, OLS, as our gold standard for measuring accuracy. While our data were constructed to satisfy the assumptions of OLS (with the exception of our outlier experiment), in most real world scenarios OLS is just one tool of many, and analysts must exercise judgment when determining how to fit their data. More complex forms of regression, and more complex fits, may not have ready analogues visually, and data concerning regression by eye may not extend to cover these cases.

Our future work is focused in three areas. Firstly, we wish to examine the impact of annotations relevant to regression (such as confidence bands, error bars, and curve boxplots~\cite{mirzargar2014curve}) on regression by eye. Can sufficient information promote caution in judgments of trend, or do initial visual estimations have too much weight to be overcome? Secondly, we wish to examine the impact of different rhetorical framings and presentations on regression by eye. Language from semiotics and rhetoric can provide testable structures for how visualizations are consumed~\cite{hullman2011visualization}: these framings (and so associated predispositions or biases about the data) could similarly impact statistical judgments. Lastly, we wish to examine techniques for overcoming bias in regression by eye and similar visual estimations of statistical quantities. Cognitive and perceptual biases are difficult to overcome, and may require exploration of novel forms of visualization \cite{micallef2012assessing}.

@@

In this paper, we examine the ability of viewers of visualizations to perform regression by eye --- the visual estimation of trend in bivariate visualizations. We show that even viewers without statistical training can reliably estimate even non-linear trends in visualizations like scatter plots and linegraphs. However, area charts are subject to within the bar bias, and promote trends with lower intercepts than other bivariate visualizations. Regression by eye is also less sensitive to outliers than simple statistical regression techniques, resulting in a divergence between trends estimated by viewers, which do not give much weight to extreme outliers, and those trends calculated by simple statistical methods, can give equal weight all points in a dataset.

>>>>

Even if designers include all relevant information, the audience may lack the statistical expertise to interpret these values, or may be misled if the data violate modeling assumptions. As a further complication, the form of visual encoding may influence viewers' inferences. For instance, viewers may be more likely to consider trends with line charts, and to compare individual values with bar charts \cite{zacks1999bars}. Visual design choices can also introduce bias, such as the visual asymmetry of bars causing ``within-the-bar'' bias \cite{newman2012bar}. Designers would benefit from guidance regarding how accurately viewers make trend estimations by eye, and to what degree different visualization types might bias these estimations.

@@

\section{Experiment 1: Slope Estimation}

We designed our first experiment to examine how accurate participants were at estimating the magnitude (slope, amplitude, or curvature) of trends in bivariate visualizations. We examined a variety of common bivariate visualizations \jeff{Such as?}. In addition to linear trends, we examined more complex relationships such as quadratic and trigonometric functions.

We presented participants with a series of bivariate visualizations, who adjusted a slider to fit the perceived trend in the points. The visualizations were one of a scatter plot, line graph, or area chart (with the filled area below the line). The slider parameterized one of three types of trend (linear, quadratic, or trigonometric). For each stimulus, participants adjusted a slider that controlled the slope of a rendered trend line (or, in the case of the quadratic trends, the curvature, or the trigonometric, the positive/negative amplitude). \jeff{To many parentheticals...}

Participants saw one of each combination of 3 chart types (scatter plot, line chart, or area chart), 8 possible slopes $\beta = \pm \{0.1,0.2,0.4,0.8\}$, and 4 bandwidths of Gaussian residuals $\sigma = \{0.05,0.1,0.15,0.2\}$, for a total of 96 stimuli. We also included an additional 4 validation stimuli otherwise excluded from analysis. The type of trend (linear, quadratic, or trigonometric) was a random factor, with 32 stimuli of each factor level. \jeff{Was this factor random or consistent across subjects?}

@@

We had three hypotheses for the first experiment:
\begin{enumerate}
	\item \textbf{As the bandwidth of the residuals increased, accuracy would decrease}. Increasing the bandwidth of the residuals results in a lower correlation coefficient and higher perceived noise in the bivariate data. Prior work indicates that these related measures correspond to decreased accuracy for aggregate tasks in bivariate visualizations~\cite{albers2014task, harrison2014ranking}.
	\item \textbf{More complex relationships would result in lower accuracy}. Quadratic and trigonometric relationships are visually more complex than linear relationships, and often require more complex statistical methods to analyze. We anticipated that estimation of these less familiar relationships would therefore be more difficult than the linear case.
	\item \textbf{Estimations would be unbiased}. That is, there would be no systematic over- or underestimation of trends.
\end{enumerate}

\subsection{Results}
We performed a three-way analysis of covariance (ANCOVA) of the effect of residual bandwidth, graph type and trend type on unsigned error \jeff{absolute error?} in estimation of trend lines. We included participant ids and the actual slope of the trend line as covariates.

Our results support our first hypothesis: \textbf{larger residuals reduce accuracy at regression by eye}.  We observed a significant main effect for the bandwidth of the Gaussian used to generate residuals ($F(1,4554)=950$, \jeff{p-value?}). The trimmed mean \jeff{trimmed how?} of absolute error increased monotonically with this bandwidth, from $7\%$ of the actual slope of the trend line when the bandwidth was $0.05$, to $32\%$ when the bandwidth was $0.20$. Figure \ref{fig:sigma} illustrates these results.

Our results fail to support our second hypothesis: \textbf{there was no statistically significant difference in estimation accuracy among linear, quadratic, or trigonometric trends}. Fit type had only a marginal main effect ($F(2,4554)=2.60$, $p=0.074$), and a post-hoc tests (using Tukey's Honest Significant Difference) did not identify any significant pairwise-interactions. The trimmed mean of absolute estimation error of linear fits was $17\%$ of the actual slope, in the middle of the $16\%$ error for quadratic fits, and $18\%$ for trignometric fits, providing evidence that the relative unfamiliarity of non-linear trends does not have a notable impact on performance.

Our results support our third hypothesis: \textbf{there was no statistically significant bias in estimations}. Participants saw an equal number of stimuli with positive and negative trend lines. If estimates of these trends were unbiased, we would expect the average signed error to be close to zero. The average signed error was $0.0008$, far less than the fidelity of the slider used to input guesses ($\Delta=0.01$). A Student's T-test failed to support the hypothesis that $mu_{error}\ne0$ ($p=0.70$) \jeff{Test statistic and degrees of freedom?}.

\section{Experiment 2: ``Within-the-Bar'' Bias}

``Within the bar'' bias is a known perceptual bias involving bar charts, in which points contained in the visual area of the glyph of the bar are deemed likelier than points outside of the glyph. Newman \& Scholl~\cite{newman2012bar} encountered this bias for a sampling task: ``how likely is this point to have been drawn from the distribution represented by this bar?'' Correll \& Gleicher~\cite{correll2014error} likewise encountered this bias for inferential tasks: ``how likely is the population mean to take a particular value, given the sample represented by this bar?''

We hypothesized that this bias would likewise occur for regression tasks when using visually asymmetric visualizations such as area charts. The slope estimation task in the previous experiment would not capture this bias, as there is no method for participants to indicate a \emph{uniform} underestimation in trends; decreasing the slope would cause underestimation at the beginning of the plot but not the end, and vice versa. We therefore designed this experiment to elicit estimates of the y-intercepts of trends. A within-the-bar bias would then appear as systematic underestimation of intercept in area charts.

@@

We had one hypothesis for the second experiment:
\begin{enumerate}
	\item \textbf{Area charts would be subject to within the bar bias.} That is, participants would estimate lower values of y-intercepts of trends than with line charts and scatter plots.
\end{enumerate}

\subsection{Results}
We performed a one-way ANCOVA of the effect of graph type on \emph{signed} error. We included residual bandwidth as a covariate, and participant ids as a random factor.

Our results support our first hypothesis: \textbf{participants systematically overestimated the intercept of trends in area charts, but not in scatter plots or line graphs}. We observed a significant main effect of graph type on signed error ($F(2,2431)=27$, $p<0.001$). A post-hoc Tukey's HSD confirmed significant differences in error between the area chart and the other two chart types, but not between scatter plots and line graphs. The trimmed mean \jeff{trimmed how?} of the signed error of estimations made with area charts was an underestimation of $-0.02$, compared to a trimmed mean of $0.002$ for the other two conditions. This underestimation corresponded to unsigned errors more than twice as large in area charts (trimmed mean of $0.04$ for intercepts that ranged from $\{-0.25,0.25\}$, compared to $0.02$ for line graphs and scatter plots). Figure \ref{fig:outlier} illustrates this result.

\section{Experiment 3. Estimation involving Outliers}

\outlierFig
OLS regression operates under the assumption that there is a unimodal, symmetric distribution of residuals surrounding the line of best fit. Extreme outliers violate this assumption, and can result in trend lines that are substantially different from those produced by more robust methods. Visual inspection can accurately identify certain classes of outliers \cite{albers2014task}. Therefore, regression by eye may afford the estimation of both outlier-robust and outlier-sensitive trends. However, cognitive biases such as anchoring (the tendency to overweight the first set of information), availability (the tendency to overweight more recent or extreme information), and the hot-hand fallacy (the tendency to assume that trends of high or low values with continue) can impact how viewers categorize and utilize outliers in their estimations \jeff{Need citations in this sentence?}. These biases can impact forecasting tasks~\cite{campbell2009anchoring, ji2001culture}. That is, regression by eye may not uniformly weight outliers, depending on their position in the plot.

@@

We had three hypotheses for the third experiment:
\begin{enumerate}
	\item \textbf{Participant estimation would be closer to a robust trend line ignoring outliers, than the non-robust OLS trend line.} That is, we assumed that in general, participants would ignore or downweight outliers when performing regression by eye.
	\item \textbf{As the number of outliers increased, estimations would be closer to the non-robust OLS fit.} We speculated on the existence of a ``tipping point'' of outlier density, beyond which participants would avoid robust fits, and interpolate between the robust and OLS fits.
	\item \textbf{Outliers at the end of the chart would result in estimations closer to the OLS fit than outliers in other locations on the plot.} Prior work on cognitive biases in forecasting tasks suggests that viewers may give more credence to more recent outliers (as indicative as new anchors points, or of an emerging ``streak'') \jeff{cite?}. We therefore believed that these points would be more heavily weighted, and the resulting estimation closer to the outlier-sensitive OLS line, than outliers ``earlier'' in the plot.
\end{enumerate}

\subsection{Results}

Our results support our first hypothesis. \textbf{Participant estimates were closer to the robust trend line than the trend line that included outliers}. Excluding conditions with zero outliers (and so the two trend lines would be identical), the trimmed mean \jeff{trimmed how?} of unsigned error was over $4$ times higher when calculated as a comparison to the OLS trend line ($0.36$) as opposed to the robust trend line ($0.08$). A Student's T-test confirmed that the error as defined by the OLS line was significantly higher than the robust line ($p<0.001$) \jeff{test statistic, dof}. However, a Student's T-test found a significant difference in unsigned error when compared to the robust trend line between estimates where no outliers were present, and those with any amount of outliers ($\mu=0.07$ vs. $\mu=0.08$, $p<0.001$). This indicates that participants are not entirely ignoring outliers, although the small effect size suggestes that participants are giving outliers significantly less weight than OLS regression.

\jeff{Whither the other hypotheses?}

@@

In many cases, designers do not explicitly encode regression information. In other cases, viewers may not have the statistical or graphical expertise to interpret such information, even when it is supplied. Yet, our results point to regression by eye as a robust and accurate method for estimating trends in bivariate data. Participants from a variety of backgrounds and levels of self-reported graphical and statistical expertise were capable of estimating both the slope and intercept of both linear and non-linear trends. That is, viewers of visualization are largely trustworthy when estimating the relationship between two variables in a plot.
\jeff{This point was not so clear in the results discussion of experiment 1. We should discuss (and visualize) the range of accuracies observed there.}

However, this general accuracy of trend estimation is not universal. Area charts are visually asymmetrical: the area below the line is filled in, and the area above it is not. We found that this asymmetry creates a within the bar bias: a systematic underestimation of the intercept of trends, due to the perceived higher likelihood of points in the filled-in area. Designers hoping to rely on regression by eye should avoid such asymmetries in their bivariate visualizations.

Likewise, viewers do not give the same weight to outliers as OLS regression, but they do not ignore them either. For noisy data, this robustness may be beneficial: people can be relied on to perform filtering operations without explicit guidance. However, in other cases, this insensitivity may result in visual estimations at odds with the statistical modeling underpinning other areas of analysis of a particular dataset \jeff{What are you trying to say here?}. The emergence of new modes and the shifting of means may not result in a subsequent shift in beliefs about trends in data. \jeff{???}

@@

Our experimental setup was intentionally simple, affording only a single free parameter for each experiment. In actual regression by eye, the viewer may simultaneously engage in multiple types of estimation: choosing a particular type of fit, ignoring outliers, and estimating the parameters of the chosen model. Errors in any one of these steps could compound, resulting in performance worse than our measures, where many of these decisions are fixed.

@@

Finally, we focused on a simple form of regression, OLS, as our standard for measuring accuracy. While our data were constructed to satisfy the assumptions of OLS (with the intentional exception of our outlier experiment), in most real world scenarios OLS is just one tool of many, and analysts must exercise judgment when determining how to fit their data. More complex models may not have ready visual analogues, and data concerning regression by eye may not extend to cover these cases.

Our future work is focused in three areas. First, we wish to examine the impact of annotations relevant to regression (such as confidence bands, error bars, and curve boxplots~\cite{mirzargar2014curve}) on regression by eye. Can sufficient information promote caution in judgments of trend, or do initial visual estimations have too much weight to be overcome? Second, we wish to examine the impact of different rhetorical framings and presentations on regression by eye. Language from semiotics and rhetoric can provide testable structures for how visualizations are consumed~\cite{hullman2011visualization}: these framings (and so associated predispositions or biases about the data) could similarly impact statistical judgments. \jeff{Can you clarify this last part? And/or provide a concrete example?} Lastly, we wish to examine techniques for overcoming bias in regression by eye and similar visual estimations of statistical quantities. Cognitive and perceptual biases are difficult to overcome, and may require exploration of novel forms of visualization~\cite{micallef2012assessing}.

@@

In this paper, we examine the ability of viewers of visualizations to perform \emph{regression by eye}\,---\,the visual estimation of trends in bivariate visualizations. We show that even viewers without statistical training can reliably estimate both linear and non-linear trends in visualizations such as scatter plots and line graphs. However, area charts are subject to a ``within the bar'' bias, leading to estimated trends with lower intercepts than other bivariate visualizations. Regression by eye is also less sensitive to outliers than simple statistical regression techniques, resulting in a divergence between trends estimated by viewers, which do not give much weight to extreme outliers, and those trends calculated by simple statistical methods, can give equal weight all points in a dataset. \jeff{Run-on sentence. Also OLS does \emph{not} give equal weight to all points, such is the nature of \textbf{squared} error.}
